---
title: "R Notebook"
output: html_notebook
---
---
NSS eggheads: Network analysis between medical providers
project: hop_team
---

Load libraries
```{r}
library(tidyverse)
library(ggplot2)
library(data.table)
library(dplyr)
library(plotly)
library(igraph)
library(igraphdata)
library(visNetwork)
```

Load in datasets

```{r}

hop_referral <- read_csv("data/provider_network_nash_cbsa.csv")

patients_community_id <- read_csv("data/neo4j_query_table_data_patients.csv")

#transactions_community_id <- read_csv("data/neo4j_query_table_data_transactions.csv")

```

Merge community ID from neo4J algorithm to the transaction dataset
```{r}

hop_referral_merged <- merge(
  hop_referral,
  
  patients_community_id,
  by.x = "from_npi", by.y = "npi"
)

hop_referral_merged  
```
rename columns
```{r}

hop_referral_merged_pcp <- rename(hop_referral_merged, c("pcp_communityid" = "communityId"))

hop_referral_merged_pcp

```

```{r}

hop_referral_merged_pcp_hosp <- merge(
  hop_referral_merged_pcp,
  patients_community_id,
  by.x = "to_npi", by.y = "npi"
)

hop_referral_merged_pcp_hosp

```

```{r}

hop_referral_pcp_hosp <- rename(hop_referral_merged_pcp_hosp, c("hospital_communityid" = "communityId"))

hop_referral_pcp_hosp

```
Save the new csv file
```{r}
write.csv(hop_referral_pcp_hosp, file = "data/hop_referral_pcp_hosp.csv")
```


Explore unique community IDs

```{r}

hop_referral_pcp_hosp %>% 
  group_by(to_organization) %>% 
  summarise(num_id = n_distinct(hospital_communityid)) %>% 
  arrange(desc(num_id))
```

Identify only VUMC network ID
```{r}
vumc_id<- hop_referral_pcp_hosp %>% 
  filter(
    to_organization == "VANDERBILT UNIVERSITY MEDICAL CENTER"
  ) %>% 
  distinct(
    hospital_communityid
  ) |> pull()
vumc_id
```

Now load the list for the Other three big hospital chains
HCA HEALTH SERVICES OF TENNESSEE, INC.

```{r}

HCA_id<- hop_referral_pcp_hosp %>% 
  filter(
    to_organization == "HCA HEALTH SERVICES OF TENNESSEE, INC."
  ) %>% 
  distinct(
    hospital_communityid
  ) |> pull()
HCA_id
```
SAINT THOMAS WEST HOSPITAL

```{r}

ST_id<- hop_referral_pcp_hosp %>% 
  filter(
    to_organization == "SAINT THOMAS WEST HOSPITAL"
  ) %>% 
  distinct(
    hospital_communityid
  ) |> pull()
ST_id
```
Now identify all other unique community id from hospitals 

```{r}
exclude_chain_id <- c(vumc_id, HCA_id, ST_id)
exclude_chain_id
#create a list of the community ids that correspond to other hospitals
single_community_id <- hop_referral_pcp_hosp |> filter(!(hospital_communityid %in% exclude_chain_id)) |> distinct(hospital_communityid) |> pull()
single_community_id
```

Create unique names for the cluster id to make grouping easier 
```{r}
#pull out community id that is exclusively vumc
only_vumc <- setdiff(vumc_id, c(HCA_id, ST_id))
only_vumc

#pull out community id that is exclusively hca
only_HCA <- setdiff(HCA_id, c(vumc_id, ST_id))
only_HCA

#pull out community id that is exclusively ST
only_ST <- setdiff(ST_id, c(vumc_id, HCA_id))
only_ST

#pull out the community id that is SHARED between VUMC and any other hospital
vumc_overlap <- setdiff(exclude_chain_id, c(only_vumc, only_HCA, only_ST))
vumc_overlap

#create a new dataframe with a case_when to get unique cluster groups for easier graphing 

hop_referral_pcp_hosp <-hop_referral_pcp_hosp |> 
  mutate(group_cluster = case_when( 
    hospital_communityid %in% only_vumc ~ 'vumc', 
    hospital_communityid %in% only_HCA ~ 'HCA', 
    hospital_communityid %in% only_ST ~ 'St Thomas', 
    hospital_communityid %in% vumc_overlap ~ 'VUMC overlap', 
    hospital_communityid %in% single_community_id[1] ~ 'community 1', 
    hospital_communityid %in% single_community_id[2] ~ 'community 2',
    hospital_communityid %in% single_community_id[3] ~ 'community 3',
    hospital_communityid %in% single_community_id[4] ~ 'community 4',
    hospital_communityid %in% single_community_id[5] ~ 'community 5',
    hospital_communityid %in% single_community_id[6] ~ 'community 6',
    hospital_communityid %in% single_community_id[7] ~ 'community 7', 
    .default= 'other'))
hop_referral_pcp_hosp
```
Create a list with more descriptive names for the community IDs for graphing
```{r}

name_for_commuityid<- c("VUMC" = only_vumc, 
  "VUMC shared"= vumc_overlap, 
  "HCA" = only_HCA, 
  "St thomas" =only_ST, 
  "community 1" = single_community_id[1], 
  "community 2" = single_community_id[2], 
  "community 3" = single_community_id[3], 
  "community 4" = single_community_id[4],
  "community 5" = single_community_id[5],
  "community 6" = single_community_id[6],
  "community 7" = single_community_id[7])
name_for_commuityid


#switch the order 
name_for_commuityid <- setNames(names(name_for_commuityid), name_for_commuityid)
name_for_commuityid
```

SM edits: Making a network graph visual

source of code: 
https://www.statworx.com/en/content-hub/blog/interactive-network-visualization-with-r#:~:text=visNetwork%20has%20many%20adjustments%20to,them%20in%20the%20right%20format.

https://cran.r-project.org/web/packages/visNetwork/vignettes/Introduction-to-visNetwork.html

https://kateto.net/network-visualization

Dataset details:
* nodes -- data frame with the nodes (aka characters) --> this will be npi
* edges -- data frame with the connections --> this will be num of patients


### Decription from code source
Nodes:

label: A column that defines how a node is labelled
value: Defines the size of a node inside the network
group: Assigns a node to a group; this can be a result of a cluster analysis or a community detection
shape: Defines how a node is presented. For example as a circle, square or triangle
color: Defines the color of a node
title: Sets the tooltip, which occurs when you hover over a node (this can be HTML or character)
shadow: Defines if a node has a shadow or not (TRUE/FALSE)
Edges:

label, title, shadow
length, width: Defines the length/width of an edge inside the network
arrows: Defines where to set a possible arrow on the edge
dashes: Defines if the edges should be dashed or not (TRUE/FALSE)
smooth: Smooth lines (TRUE/FALSE)


First make the edges, which will be the from_npi, to_npi (each a 'node') and the line will be patient count

```{r}

#Select only the 'to' and 'from' npi from the dataset, and the number of patients for the edges
data_for_edges <- hop_referral_merged_pcp_hosp |> 
  select(to_npi, from_npi, patient_count)
data_for_edges

```


```{r}
### Create a smaller width size by dividing all patient number by 100, which makes visual graphing easier 
clean_data_for_edges <- data_for_edges |> mutate(width_100 = round(patient_count/100))

#clean_data_for_edges <- data_for_edges |> select(from, to, width_100)
colnames(clean_data_for_edges) <- c('from', 'to', 'title', 'width')

## add in a text information to the 'title' column so that the figure has clear labeling

clean_data_for_edges<- clean_data_for_edges |> mutate(title= paste("patients: ", title))
clean_data_for_edges
```

then make the nodes themselves. This is JUST the list of the npi as 'id' and the group number ('community id') from the clustering algorithm

```{r}
data_for_nodes <- patients_community_id
colnames(data_for_nodes) <- c("id", "group")

data_for_nodes

```

Add in a column for the nodes that have the name for the nodes, using the full_name and to_organization information connected to the npi
```{r}

## filter out only the columns needed before the merge
#take out only the columns with the pcp npi and name from the dataframe
name_for_nodes <- hop_referral_pcp_hosp |> select(from_npi, full_name, group_cluster)

# merge it into the node in order to make it a 'label'
pcp_id_node <- left_join(patients_community_id, name_for_nodes, by= join_by (npi == from_npi))


## repeat but for to_npi 
name_for_nodes <- hop_referral_pcp_hosp |> select(to_npi, to_organization, group_cluster)

hosp_pcp_id_node <- left_join(pcp_id_node, name_for_nodes, by= join_by (npi == to_npi))
hosp_pcp_id_node

#merge the node values together 
hosp_pcp_id_node <- hosp_pcp_id_node |> mutate(group_cluster = coalesce(group_cluster.x, group_cluster.y))

#pull out only values needed for the graphing
clean_data_for_node <- hosp_pcp_id_node |>
  #Add a new column that pulls all the not-null values from the to_npi and from_npi merges
  mutate(label = tolower(coalesce(full_name, to_organization))) |>
  #pull out only the columns necessary for the nodes
  select(npi, communityId, label, group_cluster) |>
  #drop duplicates produced by the merge
  distinct()

clean_data_for_node


#replace the community ID with the descriptive names

clean_data_for_node <- clean_data_for_node |>
  mutate(community_name = recode(communityId, !!!name_for_commuityid))

clean_data_for_node
```

The graphing needs specific column names to match information, so update the column names before graphing 
```{r}
#pull out only the columns needed for graping
clean_data_for_node <- clean_data_for_node |> select(npi,community_name, label)
clean_data_for_node

#rename to match name formating for graphing
colnames(clean_data_for_node) <- c("id", "group", "title")
clean_data_for_node

#add a blank label column to remove the text on the graphing
clean_data_for_node <- clean_data_for_node |> mutate(label= "")

#remoce and duplicate id
clean_data_for_node <- clean_data_for_node |> distinct()

clean_data_for_node


```

Create the visual graph 
```{r}

visNetwork(clean_data_for_node, clean_data_for_edges, width="100%") |> 
  #create the graph
  visIgraphLayout() |>
  #change visual format for nodes
  visNodes(shape= 'dot') |>
  #change visual format for edges
  visEdges(color='gray') |>
  #Create a drop down menu 
  visOptions(highlightNearest = list(enabled = T, degree = 1, hover = T),
             #Sets which table variable to group based on a drop-down menu
             selectedBy = "group") |> 
  #define the starting seed so it groups the same way each time 
  visLayout(randomSeed=1) |>
  #create global for all groups
  visGroups() |>
  #add a legend 
  visLegend() |> 
  #add navigation tools 
  visInteraction (navigationButtons = TRUE) |> 
  visClusteringByGroup(groups, force= TRUE) |>

#add code to make dot clustering easier (so there is no overlap)
  visPhysics(stabilization = FALSE, solver = "barnesHut", barnesHut = list(gravitationalConstant = -10000))
```




check specialty for list 
hop_referral_pcp_hosp
```{r}
hop_referral_pcp_hosp
```


```{r}

user_choice <- hop_referral_pcp_hosp |>  distinct(pcp_class) |> pull()



user_choice[1]
```
```{r}
hop_referral_pcp_hosp |> filter(pcp_class == user_choice[1])
```

